
# Functions

```{r, message=F}
library(here)
library(tidyverse)
library(ggh4x)
library(pheatmap)

source("R/config_plotting.R")
source("R/utils.R")

theme_set(theme_print_function())


```

# Data path

```{r}
data_path <- here("Data processing and filtering")
```

# --- HeLa data ---

```{r}

# output from merge filter
load(file.path(data_path,"/MF35/LipidQuan output files/PE and PEO negative mode/Results/PAPSL_MF35.rdata"))

load(file.path(data_path,"/MF35/LipidQuan output files/PE and PEO negative mode/Results/PAPSL_MF35_supplementary.rdata"))

# Output from analysis
load(file.path(data_path,"MF35/LipidQuan output files/PE and PEO negative mode/Results/MF35_analysis_negative.rdata"))

# Output from 6_network_calculations_HeLa.R
load(file.path(data_path,"MF35/LipidQuan output files/PE and PEO negative mode/Results/MF35_clustering_negative_HeLa.rdata"))

```

from 4_Merge_filter.R:
* melt_classes
* mean_length_db_classes
* mean_length_db_all

from 5_Data_analysis.R:
* matrix_rel_to_wc

from 6_network_calculations_HeLa
* melt_species_cluster

# Fig 1

## 1b: detected lipids

```{r, fig.width=3, fig.height=1.5, dpi=600, out.width="20%"}


count_unique_species  <- melt_species %>%
  filter(!organelle %in% c("Whole cells")) %>%
  mutate(organelle = droplevels(organelle)) %>%
  group_by(organelle, species) %>%
  summarise(mean_molpct = geom_mean(molpct, na.rm = T), .groups = "drop") %>%
  pivot_wider(names_from = organelle, values_from = mean_molpct) %>%
  rowwise() %>%
  mutate(
    v = list(coalesce(c_across(-species), 0)),
    all_organelles    = all(v > 0),
    unique_completely = sum(v > 0) == 1,
    unique            = max(v) > sort(v, decreasing = TRUE)[2] * 10
  ) %>%
  select(-v)

count_unique_species%>%
  group_by(species) %>%
  summarise(unique = sum(unique)>0, 
            all = sum(all_organelles) >0, 
            unique_completely = sum(unique_completely)>0) %>%
  summarise(unique_enriched = sum(unique),
            unique_one_compartment = sum(unique_completely), 
            in_all_organelles = sum(all),
            total = n()) %>%
  pivot_longer(cols = everything()) %>%
  mutate(name = factor(name, levels = c("unique_enriched","unique_one_compartment",
                                        "in_all_organelles","total")))%>%
    ggplot(aes(x = name, y = -value))+
    geom_bar(stat = "identity", fill = "black", width = .16)+
    geom_point(size = 6)+
    geom_text(aes(label = value), color = "white", size = 7/3)+
    coord_flip(clip = "off")+
    scale_y_continuous(breaks = c(0,-100,-200,-300), 
                       labels = c(0,100,200,300))+
    theme(axis.line.y.left = element_blank(), 
          panel.grid.major.y = element_blank(), 
          axis.text.x = element_text(angle = 0, hjust = .5), 
          axis.text.y = element_text(hjust = 0)
          )+
    labs(y = NULL, x = NULL)+
    ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                            cols = unit(2.5, "cm"))

ggsave(
  file.path(data_path, "/figures/pdf/Fig1b_count_all_enriched_unique.pdf"),
  height = 4.5,
  width = 6,
  units = "cm"
)



```


## 1c: HeLa whole cell class mol% pie chart

```{r, fig.width=1.5, fig.height=1.5, dpi=300, out.width="20%"}
class_molpct_summary <- melt_classes%>%
  group_by(class, organelle) %>%
  summarise(molpct = geom_mean(molpct_class, na.rm = T), .groups = "drop")


class_molpct_summary %>%
  filter(organelle %in% "Whole cells") %>%
  ggplot(aes(x = "", y = molpct, fill = class))+
    geom_col(show.legend = F, color = "white", linewidth = .1)+
    geom_text(aes(label = class, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
    scale_fill_manual(values = col_class)+
    labs(x = NULL, y = NULL)  +
    theme(axis.text.x = element_blank(), 
          axis.line.x.bottom = element_blank(),
          axis.line.y.left = element_blank()) +
  ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                        cols = unit(2.5, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig1c_WC_class_molpct_circle.pdf"), height = 1.5, width = 1.5)


```



## 1d: organelle class mol% pie charts

```{r, fig.width=3, fig.height=4, dpi=100, out.width="20%"}

class_molpct_summary %>%
  filter(!organelle %in% c("LMF", "Whole cells")) %>%
  ggplot(aes(x = "", y = molpct, fill = class))+
    geom_col(show.legend = F, color = "white", linewidth = .1)+
    geom_text(aes(label = class, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
    scale_fill_manual(values = col_class)+
    labs(x = NULL, y = NULL)  +
  facet_wrap(~organelle, nrow = 3)+
    theme(axis.text.x = element_blank(), 
          axis.line.x.bottom = element_blank(),
          axis.line.y.left = element_blank()) +
  ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                        cols = unit(2.5, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig1c_WC_class_molpct>_circle.pdf"), height = 4, width = 3)

```



## 1e: all classes mol% barplot

```{r, fig.width=8, fig.height=3, dpi=600, out.width="20%"}
melt_classes %>%
  filter(!organelle %in% c( "LMF", "Peroxisomes")) %>%
  ggplot(aes(x = organelle, y = molpct_class, fill = organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", position = "dodge", show.legend = F, linewidth = 0.2, color = "black")+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_organelle)+
  geom_point(size = 0.2, show.legend = F)+
  coord_cartesian(clip = "off")+
  labs(y = "mol%", x = NULL)+
  facet_wrap(~class, nrow = 3, scales = "free_y")+
  scale_y_continuous(expand = expansion(mult = c(0,0.05)))+
  ggh4x::force_panelsizes(rows = unit(1.064,"cm"), 
                          cols = unit(1.509,"cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig1e_allClassesMolpct.pdf"), height = 4, width = 11)
```


## 1f: organelle Pearson cor heatmap

```{r}


matrix_rel_to_wc <- melt_species %>%
  filter(!organelle == "Whole cells") %>%
  mutate(organelle_exp = paste(organelle, exp, sep = "@"))%>%
  pivot_wider(names_from = organelle_exp, values_from = molpct_rel_to_wc, id_cols = species) %>%
  column_to_rownames(var = "species")

# remove NA rows
matrix_rel_to_wc <- matrix_rel_to_wc[rowSums(!is.na(matrix_rel_to_wc))>0,]
  
# log2 transform
matrix_rel_to_wc <- log2(matrix_rel_to_wc)

# Annotation matrix
cor_annotation <- data.frame(organelle = gsub("@.*", "", colnames(matrix_rel_to_wc)))
rownames(cor_annotation) <- colnames(matrix_rel_to_wc)

heatmap_pearson <- matrix_rel_to_wc %>%
  cor(use = "pairwise.complete.obs") %>%
  pheatmap::pheatmap(color = color2, 
           breaks = seq(-1,1, length.out = 101), 
           border_color = NA, 
           annotation_row = cor_annotation, 
           annotation_col = cor_annotation, 
           annotation_colors = list(organelle = col_organelle[names(col_organelle) %in% cor_annotation$organelle]),
           show_rownames = F,
           show_colnames = F,
           cellheight = 7, 
           cellwidth = 7, 
           filename = file.path(data_path,"/figures/pdf/Fig1f_rel_to_wc_cor_cluster.pdf"),
           height = 8, 
           width = 8
           )

```


## 1g: organelle PCA

```{r, fig.width=2, fig.height=2, dpi=300, out.width="50%"}
pca_df %>%
  ggplot(aes(y = -`PC1 (28.2%)`, x = `PC2 (20.7%)`, fill = organelle))+
  geom_point(size = 2, shape = 21, show.legend = F)+
  scale_fill_manual(values = col_organelle)+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(), 
        axis.ticks = element_line(size = 0.2))+
  ggh4x::force_panelsizes(rows = unit(2.5, "cm"), cols =  unit(2.5, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig1g_PCA_1.pdf"), height = 2, width = 2)

pca_df %>%
  ggplot(aes(y = -`PC3 (8.6%)`, x = `PC2 (20.7%)`, fill = organelle))+
  geom_point(size = 2, shape = 21, show.legend = F)+
  scale_fill_manual(values = col_organelle)+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(), 
        axis.ticks = element_line(size = 0.2))+
  ggh4x::force_panelsizes(rows = unit(2.5, "cm"), cols =  unit(2.5, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig1g_PCA_2.pdf"), height = 2, width = 2)
```


## 1h: organelle category mol% stacked bar plots 

```{r, fig.width=2.3, fig.height=2.1, dpi=200, out.width="20%"}

melt_species%>%
  mutate(cat = factor(cat, levels = rev(c("ST", "SL", "LGPL", "LGPLO-", "GPLO-", "GPL"))))%>%
  filter(!organelle %in% c( "Whole cells", "LMF")) %>%
  group_by(cat, organelle, species) %>%
  summarise(molpct_mean = mean(ifelse(is.na(molpct),0, molpct), na.rm = T), .groups = "drop") %>%
  group_by(cat, organelle) %>%
  summarise(cat.molpct = sum(molpct_mean, na.rm = T), .groups = "drop") %>%
  ggplot(aes(x = organelle, y = cat.molpct, fill = cat))+
    geom_bar(stat = "identity", show.legend = T, linewidth =0.1, color = "white")+
    scale_fill_manual(values = col_cat)+
    labs(y = "mol%", x = NULL)+
    scale_y_continuous(expand = expansion(0,0))+
    force_panelsizes(rows = unit(2.4, "cm"), cols =  unit(2.26, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig1h_Categories_fill_bars.pdf"), height = 2.1, width = 2.3)

```


## 1i: PG/BMP, shotgun vs LC-MS

```{r, fig.width=2.5, fig.height=1.5, dpi=300, out.width="20%"}
# Read file
lcms_data <- read.csv(file.path(data_path,"/BMP and PG analysis_LC_MS.csv"))

# Remove space at the end of species names
lcms_data$species <- str_remove(lcms_data$species, " $") 

# Add PG species that are "missing"
missing_PG <- str_replace(lcms_data$species[!str_replace(lcms_data$species, pattern = "BMP", replacement = "PG") %in% lcms_data$species], pattern = "BMP", replacement = "PG")

lcms_data <- rbind(lcms_data, data.frame(species = missing_PG, classpct = 0, rep = NA))

# Add lipid info
lcms_data <- cbind(lcms_data, lipid_info(lcms_data$species))

lcms_data$species_size <- paste(lcms_data$length, lcms_data$db, sep = ":")

lcms_data %>%
  mutate(length_color = ifelse(length < 36, "34-", ifelse(length > 38, "40+", "36,38")), 
       db_alpha = ifelse(db == 0, "0", ifelse(db == 1, "1", "2+"))) %>%
  group_by(species_size, class, length, length_color, db_alpha)%>%
  summarise(classpct = geom_mean(classpct), .groups = "drop")%>%
  ggplot(aes(x = "", y = classpct, group = species_size))+
    geom_col(show.legend = F, color = "white", linewidth = 0.1, aes(fill = as.character(length_color), alpha = db_alpha))+
    geom_text(aes(label = species_size, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
  scale_fill_brewer(type = "qual", palette = 6)+
  scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2+" = 0.4))+  
  scale_y_continuous(expand = c(0,0))+
  facet_wrap(~class)+
  ggtitle("Whole-cell HeLa BMP and PG quantification by LC-MS" )+
  labs(x = NULL, y = NULL)+
  theme(axis.text.x = element_blank(), 
      axis.line.x.bottom = element_blank(),
      axis.line.y.left = element_blank(), 
      panel.grid.major.y = element_blank()) +
  ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                          cols = unit(2.5, "cm"))

ggsave(file.path(data_path, "/Figures/pdf/Fig1i_BMP_PG_LCMS.pdf"), height = 1.5, width = 2.5)

melt_species$species_size <- paste(melt_species$length, melt_species$db, sep = ":")

melt_species %>%
  mutate(length_color = ifelse(length < 36, "34-", ifelse(length > 38, "40+", "36,38")), 
       db_alpha = ifelse(db == 0, "0", ifelse(db == 1, "1", "2+"))) %>%
  filter(class %in% "PG",
         organelle %in% c("Lysosomes", "Mitochondria"))%>%
  mutate(classpct = if_else(is.na(classpct),true = 0, false = classpct)) %>%
  group_by(species_size, length, organelle, length_color, db_alpha) %>%
  summarise(classpct = sum(classpct/n, na.rm = T), .groups = "drop") %>%
  ggplot(aes(x = "", y = classpct, group = species_size)) +
    geom_col(show.legend = F, color = "white", linewidth = 0.05, 
             aes(fill = as.character(length_color), alpha = db_alpha))+
    geom_text(aes(label = species_size, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
    scale_fill_brewer(type = "qual", palette = 6)+
    scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2+" = 0.4))+  
    facet_wrap(~organelle)+
    ggtitle("HeLa Mitochondria vs Lysosomes PG/BMP" )+
    labs(x = NULL, y = NULL)+
    theme(axis.text.x = element_blank(), 
        axis.line.x.bottom = element_blank(),
        axis.line.y.left = element_blank(), 
        panel.grid.major.y = element_blank()) +
    ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                            cols = unit(2.5, "cm"))

ggsave(file.path(data_path, "/Figures/pdf/Fig1i_mito_lyso.pdf"), height = 1.5, width = 2.5)


```


# Fig 2

## 2a: heatmap relative to whole cells

```{r}
class_order <- names(col_class)

order_species <- melt_species %>%
  mutate(class = factor(class, levels = class_order))%>%
  arrange(class, length, db) %>%
  pull(species) %>%
  unique()

order_df <- data.frame(order_species, class = lipid_class_name(order_species)) %>%
  column_to_rownames(var = "order_species")

melt_species %>%
  filter(!organelle =="Whole cells")%>%
  dplyr::mutate(exp_organelle = paste(exp, organelle, sep = "_"), 
                molpct_rel_to_wc = log2(molpct_rel_to_wc)) %>%
  arrange(organelle, exp) %>%
  dplyr::select(exp_organelle, species, molpct_rel_to_wc) %>%
  mutate(species = factor(species, levels = order_species[order_species %in% species])) %>%
  pivot_wider(names_from = exp_organelle, values_from = molpct_rel_to_wc) %>%
  arrange(species) %>%
  column_to_rownames("species") %>%
  pheatmap::pheatmap(cluster_cols = F, 
                   cluster_rows = F,
                   color = color2, 
                   breaks = seq(-2,2,length.out = 101), 
                   na_col = "dark blue", 
                   filename = file.path(data_path,"/figures/pdf/Fig2a_heatmap_rel_to_WC.pdf"),
                   height = 10,
                   width = 7, 
                   cellwidth = 3,
                   border_color = NA,          
                   annotation_row = order_df["class"],
                   annotation_colors = list(class = col_class[names(col_class) %in% order_df$class]))

```

2b: Network

## 2c: Line plots with network

```{r, fig.width=1.7, fig.height=5.5, dpi=100, out.width="20%"}

melt_species_cluster <- clusters %>%
  dplyr::select(species, cluster_name)%>%
inner_join(melt_species, ., by = join_by(species)) %>%
  filter(!organelle %in% "Whole cells") %>%
  group_by(exp, organelle) %>%
  mutate(sum_molpct = sum(molpct, na.rm = T), 
         molpct_noChol = molpct/sum_molpct*100) %>%
  group_by(species) %>%
  mutate(max_molpct = max(molpct_noChol, na.rm = T), 
         norm_molpct_noChol = molpct_noChol/max_molpct) %>%
  group_by(organelle, species) %>%
  mutate(n = n()) %>%
  # Average including 'undetected' lipids as 0
  group_by(organelle) %>%
  mutate(n = max(n), 
         norm_molpct_noChol = norm_molpct_noChol / n) %>%
  group_by(organelle, species) %>%
  mutate(norm_molpct_noChol = sum(norm_molpct_noChol))

cluster_colors <- c("I" = "#97bcd7", 
                                "II" = "#f574a4",
                                "III" = "#aa2451",
                                "IV" = "#7a183c",
                                "V" = "#243f5a",
                                "VI" = "#358b59",
                                "VII" = "grey",
                                "VIII" = "grey",
                                "IX" = "grey")

# Plot without cholesterol
melt_species_cluster %>%
  filter(!cluster_name %in% c("VII", "VIII", "IX"))%>%
  ggplot(aes(x = organelle, y = norm_molpct_noChol, group = species))+
  facet_wrap(~cluster_name, scales = "fixed", ncol = 1)+
  geom_line(stat = "identity", fun = "mean", linewidth = 0.2, color = "grey")+
  geom_line(aes(group = cluster_name, color = cluster_name), 
            stat = "summary", fun = "mean", linewidth = 2)+
  scale_color_manual(values = cluster_colors)+
  labs(y = "Relative enrichment", x = NULL, color = "Cluster")+
  theme(panel.grid.major.x = element_line(linewidth = .2, color = "grey", linetype = 2),
        panel.grid.major.y = element_blank())+
  ggh4x::force_panelsizes(cols = unit(1.94,"cm"),
                          rows = unit(1.33,"cm"))

ggsave(file.path(data_path, "/Figures/pdf/Fig2c_species_cluster_lines_without_chol.pdf"), height = 5.5, width = 1.7)

```


## 2d: Dot plots - species and clusters

```{r, fig.width=6, fig.height=1.5, dpi=300, out.width="20%"}

melt_species_cluster%>%
  filter(cat %in% "GPL",
         class != "CL") %>%
ggplot(aes(x = length, y = db, fill = cluster_name))+
  geom_point(shape = 21, size = 1.8, stroke = 0.2)+
  facet_wrap(~class, nrow = 1)+
  scale_fill_manual(values = cluster_colors)+
  coord_cartesian(clip = "off")+
  scale_x_continuous(breaks = seq(28,80,2))+
  scale_y_continuous(breaks = seq(0,12,1), limits = c(0,12))+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(linewidth = .2, color = "grey", linetype = 2))+
  ggh4x::force_panelsizes(cols = unit(1.38,"cm"),
                          rows = unit(2,"cm"))
  
ggsave(file.path(data_path, "/Figures/pdf/Fig2d_Species_bubles_cluster_GPL.pdf"), height = 1.5, width = 6)


melt_species_cluster%>%
  filter(cat %in% "GPL",
         class == "CL") %>%
ggplot(aes(x = length, y = db, fill = cluster_name))+
  geom_point(shape = 21, size = 1.8, stroke = 0.2)+
  facet_wrap(~class, nrow = 1)+
  scale_fill_manual(values = cluster_colors)+
  coord_cartesian(clip = "off")+
  scale_x_continuous(breaks = seq(28,80,2))+
  scale_y_continuous(breaks = seq(0,12,1), limits = c(0,12))+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(linewidth = .2, color = "grey", linetype = 2))+
  ggh4x::force_panelsizes(cols = unit(2*6/13,"cm"),
                          rows = unit(2,"cm"))
  
ggsave(file.path(data_path, "/Figures/pdf/Fig2d_Species_bubles_cluster_CL.pdf"), height = 1.5, width = 2)

melt_species_cluster%>%
  filter(cat %in% "GPLO-") %>%
ggplot(aes(x = length, y = db, fill = cluster_name))+
  geom_point(shape = 21, size = 1.8, stroke = 0.2)+
  facet_wrap(~class, nrow = 1)+
  scale_fill_manual(values = cluster_colors)+
  coord_cartesian(clip = "off")+
  scale_x_continuous(breaks = seq(28,80,2))+
  scale_y_continuous(breaks = seq(0,12,1), limits = c(0,12))+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(linewidth = .2, color = "grey", linetype = 2))+
  ggh4x::force_panelsizes(cols = unit(1.38,"cm"),
                          rows = unit(2,"cm"))
  
ggsave(file.path(data_path, "/Figures/pdf/Fig2d_Species_bubles_cluster_GPLO.pdf"), height = 1.5, width = 4)


melt_species_cluster%>%
  filter(cat %in% c("LGPL", "LGPLO-")) %>%
ggplot(aes(x = length, y = db, fill = cluster_name))+
  geom_point(shape = 21, size = 1.8, stroke = 0.2)+
  facet_wrap(~class, nrow = 1)+
  scale_fill_manual(values = cluster_colors)+
  coord_cartesian(clip = "off")+
  scale_x_continuous(breaks = seq(10,28,2))+
  scale_y_continuous(breaks = seq(0,12,1), limits = c(0,12))+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(linewidth = .2, color = "grey", linetype = 2))+
  ggh4x::force_panelsizes(cols = unit(5/13*2,"cm"),
                          rows = unit(2,"cm"))
  
ggsave(file.path(data_path, "/Figures/pdf/Fig2d_Species_bubles_cluster_LGPL.pdf"), height = 1.5, width = 5)


melt_species_cluster%>%
  filter(cat %in% "SL") %>%
ggplot(aes(x = length, y = db, fill = cluster_name))+
  geom_point(shape = 21, size = 1.8, stroke = 0.2)+
  facet_wrap(~class, nrow = 1)+
  scale_fill_manual(values = cluster_colors)+
  coord_cartesian(clip = "off")+
  scale_x_continuous(breaks = seq(28,50,2))+
  scale_y_continuous(breaks = seq(0,12,1), limits = c(0,3))+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(linewidth = .2, color = "grey", linetype = 2))+
  ggh4x::force_panelsizes(cols = unit(7/13*2,"cm"),
                          rows = unit(0.53,"cm"))
  
ggsave(file.path(data_path, "/Figures/pdf/Fig2d_Species_bubles_cluster_SL.pdf"), height = 1.5, width = 5)
```


## 2e: Violin - inter/intra class

```{r, fig.width=1, fig.height=3, dpi=300, out.width="20%"}
cor_table%>%
ggplot(aes(x = same_classlabel, y = cor, fill = same_classlabel))+
  geom_violin(quantile.linetype = 1, show.legend = F, linewidth = 0.2)+
  scale_fill_manual(values = c("dark grey", "light grey"))+
  labs(fill = NULL, x = NULL)+
  labs(y = "Pearson's r")+
  ggh4x::force_panelsizes(cols = unit(1.3,"cm"),
                          rows = unit(2.4,"cm"))

ggsave(file.path(data_path, "/Figures/pdf/Fig2e_violin_class.pdf"), height = 3, width = 1)


```


2f Prism

2g Prism

2h Prism

## 2i: Sphingolipids stacked barplot

```{r, fig.width=2.1, fig.height=2, dpi=300, out.width="20%"}

melt_classes%>%
  dplyr::filter(!organelle %in% c("Whole cells", "LMF"), 
                cat == "SL") %>%
  dplyr::mutate(class = ifelse(class %in% c("SM", "Cer"), as.character(class), "GlycoSLs")) %>%
  dplyr::mutate(class = factor(class, levels = c("SM", "Cer", "GlycoSLs"))) %>%
  dplyr::group_by(cat, organelle, class, exp) %>%
  dplyr::summarise(molpct = sum(molpct_class, na.rm = T), .groups = "drop") %>%
  dplyr::group_by(cat, organelle, class) %>%
  dplyr::summarise(molpct_mean = mean(ifelse(is.na(molpct),0, molpct)), .groups = "drop") %>%
  
ggplot(aes(x = organelle, y = molpct_mean, fill = class))+
  geom_bar(stat = "identity", color = "white", linewidth = 0.2, show.legend = T, position = "fill")+
  scale_y_continuous(expand = expansion(0,0))+
  scale_fill_manual(values = c(col_class[names(col_class) %in% c("SM", "Cer")], "GlycoSLs" = "#EAACA6" ))+  
  labs(x = NULL, y = "% of total non-PC diacylGPLs")+
  ggh4x::force_panelsizes(cols = unit(2.1,"cm"),
                        rows = unit(2,"cm"))

ggsave(file.path(data_path, "/Figures/pdf/Fig2i_SLs.pdf"), height = 2, width = 2.1)

```

# Fig 3

3a - Scheme

3b - Prism

## 3c: PCO- / PC ratio barplot 

```{r, fig.width=1.5, fig.height=1.5, dpi=300, out.width="20%"}

ratio_PC_PCO <- melt_classes %>%
  filter(organelle != "Whole cells",
         class == "PCO-") %>%
  dplyr::select(exp, organelle, molpct_pco =molpct_class) %>%
  inner_join(melt_classes) %>%
  mutate(molpct_ratio =  molpct_pco / molpct_class) %>%
  filter(class == "PC")
  

ratio_PC_PCO%>%
ggplot(aes(x = organelle, y = molpct_ratio, fill = organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", show.legend = F)+
  geom_point(size = 0.2, show.legend = F)+
  scale_y_continuous(expand = c(0,0))+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_organelle[!names(col_organelle) %in% c("b", "blank", "Blank", "Ctrl", "Peroxisomes", "Whole cells", "LMF")])+
  coord_cartesian(clip = "off")+
  labs(x = NULL, y = "PCO-/PC ratio")+
  theme(axis.text.x = element_blank())+
  ggh4x::force_panelsizes(cols = unit(1.82,"cm"),
                        rows = unit(2.46,"cm"))


ggsave(file.path(data_path, "/Figures/pdf/Fig3b_PCO_PC_ratio.pdf"), height = 1.5, width = 1.5)
```


3d: Correlation intra/inter class - Prism

3e: Cosine similarity GPLS - Prism

## 3f: HeLa whole cells class pie

```{r, fig.width=4, fig.height=4, dpi=100, out.width="20%"}

melt_species %>%
  mutate(length_color = ifelse(length < 36, "34-", ifelse(length > 38, "40+", "36,38")), 
       db_alpha = ifelse(db == 0, "0", ifelse(db == 1, "1", "2+"))) %>%
  filter(class %in% c("PC", "PS", "PG", "PI", "PE", "PA", "PEO-", "PCO-"),
         organelle %in% c("Whole cells"))%>%
  mutate(classpct = if_else(is.na(classpct),true = 0, false = classpct)) %>%
  group_by(species_size, length, organelle, length_color, db_alpha, class) %>%
  summarise(classpct = sum(classpct/n, na.rm = T), .groups = "drop") %>%
  ggplot(aes(x = "", y = classpct, group = species_size)) +
    geom_col(show.legend = F, color = "white", linewidth = 0.05, 
             aes(fill = as.character(length_color), alpha = db_alpha))+
    geom_text(aes(label = species_size, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
    scale_fill_manual(values = c("blue", "orange", "red"))+
    scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2+" = 0.4))+  
    facet_wrap(~class)+
    labs(x = NULL, y = NULL)+
    theme(axis.text.x = element_blank(), 
        axis.line.x.bottom = element_blank(),
        axis.line.y.left = element_blank(), 
        panel.grid.major.y = element_blank()) +
    ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                            cols = unit(2.5, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig3f_WC_class_molpct_circle.pdf"), height = 4, width = 4)
```


3g - HeLa data - Prism

## 3h: PEO-/PE, PEO-/PC ratio barplots 

```{r, fig.width=1.5, fig.height=1.5, dpi=300, out.width="20%"}

ratio_PE_PEO <- melt_classes %>%
  filter(organelle != "Whole cells",
         class == "PEO-") %>%
  dplyr::select(exp, organelle, molpct_pco =molpct_class) %>%
  inner_join(melt_classes) %>%
  mutate(molpct_ratio =  molpct_pco / molpct_class) %>%
  filter(class == "PE")


ratio_PC_PEO <- melt_classes %>%
  filter(organelle != "Whole cells",
         class == "PEO-") %>%
  dplyr::select(exp, organelle, molpct_pco =molpct_class) %>%
  inner_join(melt_classes) %>%
  mutate(molpct_ratio =  molpct_pco / molpct_class) %>%
  filter(class == "PC")

ratio_PCO_PEO<- melt_classes %>%
  filter(organelle != "Whole cells",
         class == "PEO-") %>%
  dplyr::select(exp, organelle, molpct_pco =molpct_class) %>%
  inner_join(melt_classes) %>%
  mutate(molpct_ratio =  molpct_pco / molpct_class) %>%
  filter(class == "PCO-")
  

ratio_PE_PEO%>%
ggplot(aes(x = organelle, y = molpct_ratio, fill = organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", show.legend = F)+
  geom_point(size = 0.2, show.legend = F)+
  scale_y_continuous(expand = c(0,0))+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_organelle[!names(col_organelle) %in% c("b", "blank", "Blank", "Ctrl", "Peroxisomes", "Whole cells", "LMF")])+
  coord_cartesian(clip = "off")+
  labs(x = NULL, y = "PEO-/PE ratio")+
  theme(axis.text.x = element_blank())+
  ggh4x::force_panelsizes(cols = unit(1.82,"cm"),
                        rows = unit(2.46,"cm"))


ggsave(file.path(data_path, "/Figures/pdf/Fig3h_PEO_PE_ratio.pdf"), height = 1.5, width = 1.5)

ratio_PC_PEO%>%
ggplot(aes(x = organelle, y = molpct_ratio, fill = organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", show.legend = F)+
  geom_point(size = 0.2, show.legend = F)+
  scale_y_continuous(expand = c(0,0))+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_organelle[!names(col_organelle) %in% c("b", "blank", "Blank", "Ctrl", "Peroxisomes", "Whole cells", "LMF")])+
  coord_cartesian(clip = "off")+
  labs(x = NULL, y = "PEO-/PC ratio")+
  theme(axis.text.x = element_blank())+
  ggh4x::force_panelsizes(cols = unit(1.82,"cm"),
                        rows = unit(2.46,"cm"))


ggsave(file.path(data_path, "/Figures/pdf/Fig3h_PEO_PC_ratio.pdf"), height = 1.5, width = 1.5)


ratio_PCO_PEO%>%
ggplot(aes(x = organelle, y = molpct_ratio, fill = organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", show.legend = F)+
  geom_point(size = 0.2, show.legend = F)+
  scale_y_continuous(expand = c(0,0))+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_organelle[!names(col_organelle) %in% c("b", "blank", "Blank", "Ctrl", "Peroxisomes", "Whole cells", "LMF")])+
  coord_cartesian(clip = "off")+
  labs(x = NULL, y = "PEO-/PCO- ratio")+
  theme(axis.text.x = element_blank())+
  ggh4x::force_panelsizes(cols = unit(1.82,"cm"),
                        rows = unit(2.46,"cm"))


ggsave(file.path(data_path, "/Figures/pdf/Fig3h_PEO_PCO_ratio.pdf"), height = 1.5, width = 1.5)
```

## 3i: HeLa: GPL categories and ether lipids stacked bar plots

```{r, fig.width=2.3, fig.height=2.1, dpi=200, out.width="20%"}

melt_species%>%
  mutate(cat = ifelse(class %in% c("PCO-", "PEO-"), class, as.character(cat)))%>%
  filter(!organelle %in% c( "Whole cells", "LMF"), 
         cat %in% c(  "PEO-", "PCO-","LGPL", "GPL")) %>%
  mutate(cat = factor(cat, levels = c("GPL","LGPL","PCO-", "PEO-")))%>%
  group_by(cat, organelle, species) %>%
  summarise(molpct_mean = mean(ifelse(is.na(molpct),0, molpct), na.rm = T), .groups = "drop") %>%
  group_by(cat, organelle) %>%
  summarise(cat.molpct = sum(molpct_mean, na.rm = T), .groups = "drop") %>%
  group_by(organelle) %>%
  mutate(molpct_total = sum(cat.molpct), 
         cat.molpct = cat.molpct/molpct_total*100)%>%
  ggplot(aes(x = organelle, y = cat.molpct, fill = cat))+
    geom_bar(stat = "identity", show.legend = T, linewidth =0.1, color = "white")+
#    scale_fill_manual(values = col_cat)+
    labs(y = "mol% of total GPLs", x = NULL)+
    scale_y_continuous(expand = expansion(0,0))+
    force_panelsizes(rows = unit(2.4, "cm"), cols =  unit(2.26, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig3i_CategoriesEtherlipids_fill_bars.pdf"), height = 2.1, width = 2.3)
```

3j - Prism

# Fig 4

4a - Cytoscape HeLa network

4b - prism

## 4c: PC species molpct

```{r, fig.width=2, fig.height=3, dpi=200, out.width="50%"}
melt_species %>%
  filter(!organelle %in% c( "LMF", "Peroxisomes"),
         species %in% c("PC 30:0", "PC 32:0")) %>%
  ggplot(aes(x = organelle, y = molpct, fill = organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", position = "dodge", show.legend = F, linewidth = 0.2, color = "black")+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_organelle)+
  geom_point(size = 0.2, show.legend = F)+
  coord_cartesian(clip = "off")+
  labs(y = "mol%", x = NULL)+
  facet_wrap(~species, nrow = 3, scales = "free_y")+
  scale_y_continuous(expand = expansion(mult = c(0,0.05)))+
  ggh4x::force_panelsizes(rows = unit(1.064,"cm"), 
                          cols = unit(1.509,"cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig4x_PCMolpct.pdf"), height = 4, width = 11)
```


## 4d: organelle class pie charts

```{r, fig.width=5, fig.height=10, dpi=200, out.width="50%"}

melt_species %>%
  mutate(length_color = ifelse(length < 36, "34-", ifelse(length > 38, "40+", "36,38")), 
       db_alpha = ifelse(db == 0, "0", ifelse(db == 1, "1", "2+"))) %>%
  filter(class %in% c("PC", "PS", "PI", "PE", "PEO-", "PCO-"),
         organelle %in% c("Mitochondria", "ER", "Plasma membrane", "Lysosomes"))%>%
  mutate(classpct = if_else(is.na(classpct),true = 0, false = classpct)) %>%
  group_by(species_size, length, organelle, length_color, db_alpha, class) %>%
  summarise(classpct = sum(classpct/n, na.rm = T), .groups = "drop") %>%
  ggplot(aes(x = "", y = classpct, group = species_size)) +
    geom_col(show.legend = F, color = "white", linewidth = 0.05, 
             aes(fill = as.character(length_color), alpha = db_alpha))+
    geom_text(aes(label = species_size, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
    scale_fill_manual(values = c("blue", "orange", "red"))+
    scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2+" = 0.4))+  
    facet_wrap(~class+organelle, ncol = 4)+
    labs(x = NULL, y = NULL)+
    theme(axis.text.x = element_blank(), 
        axis.line.x.bottom = element_blank(),
        axis.line.y.left = element_blank(), 
        panel.grid.major.y = element_blank()) +
    ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                            cols = unit(2.5, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig4d_organelle_class_molpct_circle.pdf"), height = 10, width = 5)
```


4e: prism
4f: prism

## 4g: Mean db and length, HeLa

```{r, fig.width=5, fig.height=2.5, dpi=300, out.width="20%"}
mean_length_db_classes%>%
  filter(!organelle %in% c("LMF", "Whole cells"),
         cat %in% c( "GPL", "GPLO-" )| class %in% c("SM"),
         !class %in% c("CL", "PG", "PA")) %>%
  ggplot(aes(x = class, y = mean_db, group = organelle, fill = organelle))+
  geom_point(shape = 21,
             size = 2,
             position = position_dodge(width = 0.9),
             stat = "summary", fun = "mean") +
  stat_summary(fun.data = "mean_sdl", 
               fun.args = c(mult = 1), 
               show.legend = F,
               position = position_dodge(width = 0.9), 
               linewidth = .5, 
               geom = "linerange")+
  scale_fill_manual(values = col_organelle)+
  labs(y = "Mean no. of DBs", x = NULL)+
  ggh4x::force_panelsizes(rows = unit(3, "cm"), cols =  unit(7, "cm")) 

ggsave(file.path(data_path, "/Figures/pdf/Fig4g_length_db.pdf"), height = 2.5, width = 5)
```


## 4h: Mean db and length, all classes

```{r, fig.width=3, fig.height=2.5, dpi=300, out.width="20%"}
mean_length_db_all%>%
  filter(!organelle %in% c("LMF", "Whole cells")) %>%
  ggplot(aes(x = organelle, y = mean_db, group = organelle, fill = organelle))+
  geom_point(shape = 21,
             size = 2,
             position = position_dodge(width = 0.9),
             stat = "summary", fun = "mean") +
  stat_summary(fun.data = "mean_sdl", 
               fun.args = c(mult = 1), 
               show.legend = F,
               position = position_dodge(width = 0.9), 
               linewidth = .5, 
               geom = "linerange")+
  scale_fill_manual(values = col_organelle)+
  labs(y = "Mean no. of DBs", x = NULL)+
  ggh4x::force_panelsizes(rows = unit(3, "cm"), cols =  unit(2, "cm")) 

ggsave(file.path(data_path, "/Figures/pdf/Fig4h_db.pdf"), height = 2.5, width = 3)

mean_length_db_all%>%
  filter(!organelle %in% c("LMF", "Whole cells")) %>%
  ggplot(aes(x = organelle, y = mean_length, group = organelle, fill = organelle))+
  geom_point(shape = 21,
             size = 2,
             position = position_dodge(width = 0.9),
             stat = "summary", fun = "mean") +
  stat_summary(fun.data = "mean_sdl", 
               fun.args = c(mult = 1), 
               show.legend = F,
               position = position_dodge(width = 0.9), 
               linewidth = .5, 
               geom = "linerange")+
  scale_fill_manual(values = col_organelle)+
  labs(y = "Mean no. of DBs", x = NULL)+
  ggh4x::force_panelsizes(rows = unit(3, "cm"), cols =  unit(2, "cm")) 

ggsave(file.path(data_path, "/Figures/pdf/Fig4h_length_db.pdf"), height = 2.5, width = 3)
```


4i: Prism

## 4j: ER vs Mito PG/BMP

```{r, fig.width=2.5, fig.height=1.5, dpi=300, out.width="20%"}

melt_species %>%
  mutate(length_color = ifelse(length < 36, "34-", ifelse(length > 38, "40+", "36,38")), 
       db_alpha = ifelse(db == 0, "0", ifelse(db == 1, "1", "2+"))) %>%
  filter(class %in% "PG",
         organelle %in% c("ER", "Mitochondria"))%>%
  mutate(classpct = if_else(is.na(classpct),true = 0, false = classpct)) %>%
  group_by(species_size, length, organelle, length_color, db_alpha) %>%
  summarise(classpct = sum(classpct/n, na.rm = T), .groups = "drop") %>%
  ggplot(aes(x = "", y = classpct, group = species_size)) +
    geom_col(show.legend = F, color = "white", linewidth = 0.05, 
             aes(fill = as.character(length_color), alpha = db_alpha))+
    geom_text(aes(label = species_size, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
    scale_fill_brewer(type = "qual", palette = 6)+
    scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2+" = 0.4))+  
    facet_wrap(~organelle)+
    labs(x = NULL, y = NULL)+
    theme(axis.text.x = element_blank(), 
        axis.line.x.bottom = element_blank(),
        axis.line.y.left = element_blank(), 
        panel.grid.major.y = element_blank()) +
    ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                            cols = unit(2.5, "cm"))

ggsave(file.path(data_path, "/Figures/pdf/Fig4j_mito_ER.pdf"), height = 1.5, width = 2.5)
```


# --- HCT116 vs HKh-2 data ---

```{r}

MF36_env <- new.env()
# output from merge filter
load(file.path(data_path,"/MF36/LipidQuan output files/PE and PEO negative mode/Results/PAPSL_MF36.rdata"), envir = MF36_env)

melt_classes_MF36 <- MF36_env$melt_classes
```


# Fig 5

## 5a: whole cell pie charts molpct class

```{r, fig.width=2.5, fig.height=1.5, dpi=200, out.width="20%"}
class_molpct_summary_MF36 <- melt_classes_MF36%>%
  group_by(cell, class, organelle) %>%
  summarise(molpct = geom_mean(molpct_class, na.rm = T), .groups = "drop") %>%
  mutate(class = factor(class, levels = order_class[order_class %in% class]), 
         cell = factor(cell, levels = c("HKH2", "HCT116")))


class_molpct_summary_MF36 %>%
  filter(organelle %in% "Whole cells") %>%
  ggplot(aes(x = "", y = molpct, fill = class))+
    geom_col(show.legend = F, color = "white", linewidth = .1)+
    geom_text(aes(label = class, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
    scale_fill_manual(values = col_class)+
    labs(x = NULL, y = NULL)  +
  facet_wrap(~cell) +
    theme(axis.text.x = element_blank(), 
          axis.line.x.bottom = element_blank(),
          axis.line.y.left = element_blank()) +
  ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                        cols = unit(2.5, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig5a_WC_class_molpct_circle.pdf"), height = 1.5, width = 2.5)

```

## 5b: organelle cell pie charts molpct class

```{r, fig.width=8, fig.height=3, dpi=200, out.width="20%"}


class_molpct_summary_MF36 %>%
  filter(!organelle %in% "Whole cells") %>%
  ggplot(aes(x = "", y = molpct, fill = class))+
    geom_col(show.legend = F, color = "white", linewidth = .1)+
    geom_text(aes(label = class, x=1.5), size = 2, position = position_stack(vjust = 0.5))+
    coord_polar(theta = "y",direction = -1)+
    scale_fill_manual(values = col_class)+
    labs(x = NULL, y = NULL)  +
  facet_wrap(~cell+organelle, nrow = 2) +
    theme(axis.text.x = element_blank(), 
          axis.line.x.bottom = element_blank(),
          axis.line.y.left = element_blank()) +
  ggh4x::force_panelsizes(rows = unit(2.5, "cm"),
                        cols = unit(2.5, "cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig5b_organelle_class_molpct_circle.pdf"), height = 3, width = 8)


```

## 5c: all classes mol% barplot

```{r, fig.width=8, fig.height=3, dpi=600, out.width="20%"}

melt_classes_MF36 %>%
  filter(!organelle %in% c( "LMF", "Peroxisomes")) %>%
  mutate(cell_organelle = paste(cell, organelle, sep = "_")) %>%
  ggplot(aes(x = organelle, y = molpct_class, fill = cell_organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", position = "dodge", 
           show.legend = F, linewidth = 0.2, color = "black")+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, 
               position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_cell_organelle)+
  geom_point(size = 0.2, show.legend = F)+
  coord_cartesian(clip = "off")+
  labs(y = "mol%", x = NULL)+
  facet_wrap(~class, nrow = 3, scales = "free_y")+
  scale_y_continuous(expand = expansion(mult = c(0,0.05)))+
  ggh4x::force_panelsizes(rows = unit(1.064,"cm"), 
                          cols = unit(1.509,"cm"))

ggsave(file.path(data_path,"/figures/pdf/Fig5c_allClassesMolpct.pdf"), height = 4, width = 11)
```


## 5d: PEO-/PE, PEO-/PC ratio barplots 

```{r, fig.width=2.5, fig.height=2, dpi=300, out.width="20%"}

ratio_PE_PEO <- melt_classes_MF36 %>%
  filter(organelle != "Whole cells",
         class == "PEO-") %>%
  dplyr::select(cell, exp, organelle, molpct_r = molpct_class) %>%
  inner_join(melt_classes_MF36) %>%
  mutate(molpct_ratio =  molpct_r / molpct_class) %>%
  filter(class == "PE") %>%
  mutate(cell = factor(cell, levels = c("HKH2", "HCT116")))

ratio_PE_PEO%>%
ggplot(aes(x = organelle, y = molpct_ratio, fill = organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", show.legend = F)+
  geom_point(size = 0.2, show.legend = F)+
  scale_y_continuous(expand = c(0,0))+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_organelle[!names(col_organelle) %in% c("b", "blank", "Blank", "Ctrl", "Peroxisomes", "Whole cells", "LMF")])+
  coord_cartesian(clip = "off")+
  facet_wrap(~cell, scale = "free_y") +
  labs(x = NULL, y = "PEO-/PE ratio")+
  theme(axis.text.x = element_blank())+
  ggh4x::force_panelsizes(cols = unit(1.82,"cm"),
                        rows = unit(2.46,"cm"))

ggsave(file.path(data_path, "/Figures/pdf/Fig5d_PEO_PE_ratio.pdf"), height = 2, width = 2.5)


ratio_PC_PCO <- melt_classes_MF36 %>%
  filter(organelle != "Whole cells",
         class == "PCO-") %>%
  dplyr::select(cell, exp, organelle, molpct_r = molpct_class) %>%
  inner_join(melt_classes_MF36) %>%
  mutate(molpct_ratio =  molpct_r / molpct_class) %>%
  filter(class == "PC") %>%
  mutate(cell = factor(cell, levels = c("HKH2", "HCT116")))

ratio_PC_PCO%>%
ggplot(aes(x = organelle, y = molpct_ratio, fill = organelle))+
  geom_bar(stat = "summary", fun = "geom_mean", show.legend = F)+
  geom_point(size = 0.2, show.legend = F)+
  scale_y_continuous(expand = c(0,0))+
  stat_summary(fun.data = geom_mean_sdl, geom = "errorbar", width = 0.4, position= position_dodge(width = .9), linewidth = 0.2)+
  scale_fill_manual(values = col_organelle[!names(col_organelle) %in% c("b", "blank", "Blank", "Ctrl", "Peroxisomes", "Whole cells", "LMF")])+
  coord_cartesian(clip = "off")+
  facet_wrap(~cell, scale = "free_y") +
  labs(x = NULL, y = "PCO-/PC ratio")+
  theme(axis.text.x = element_blank())+
  ggh4x::force_panelsizes(cols = unit(1.82,"cm"),
                        rows = unit(2.46,"cm"))

ggsave(file.path(data_path, "/Figures/pdf/Fig5d_PCO_PC_ratio.pdf"), height = 2, width = 2.5)

```

## 5i: ethanolamine HCT116 vs HKH2

```{r, fig.width=5, fig.height=2.5, dpi=300, out.width="20%"}

mf38d_env <- new.env()

load(file.path(data_path,"MF38d/LipidQuan output files/PE and PEO negative mode/results/PAPSL_MF38d.rdata"),envir = mf38d_env)

melt_classes_mf38d <- mf38d_env$melt_classes

melt_classes_mf38d$cell <- factor(melt_classes_mf38d$cell, levels = c("HKH2", "HCT116"))

melt_classes_mf38d$treatment <- factor(melt_classes_mf38d$treatment, levels = c("none", "eth3"))


melt_classes_mf38d%>%
  filter(class %in% c("PE", "PEO-", "PC", "PCO-"))%>%
  ggplot(aes (x = cell, y = molpct_class, fill = treatment))+
  geom_bar(stat = "summary", fun = "geom_mean", position = "dodge")+
  stat_summary(fun.data = geom_mean_sdl, geom= "errorbar", width = .4, linewidth = 0.2, position = position_dodge(width = 0.9))+
  geom_point(size = 0.2, show.legend = F, position = position_dodge(width = 0.9))+
  scale_y_continuous(expand = expansion( mult = c(0,0.025)))+
  scale_fill_brewer(palette = 1)+
  facet_wrap(~class, scales = "free", nrow = 1)+
  labs(y = "Lipid abundance (mol%)", x = NULL)+
  ggh4x::force_panelsizes(rows = unit(3, "cm"), cols =  unit(2, "cm")) 

ggsave(file.path(data_path, "/Figures/pdf/Fig5j_EA.pdf"), height = 2.5, width = 5)
```

## 5j: ethamolamine HCT116 vs HKH2

```{r, fig.width=5, fig.height=2.5, dpi=300, out.width="20%"}

mf38c_env <- new.env()

load(file.path(data_path,"MF38c/LipidQuan output files/PE and PEO negative mode/results/PAPSL_MF38c.rdata"),envir = mf38c_env)

melt_classes_mf38c <- mf38c_env$melt_classes %>%
  mutate(cell = factor(cell, levels = c("HKH2", "HCT116")), 
         D4 = ifelse(grepl("D4", class), "D4", "no_label"), 
         class = gsub("OD4", "O-", class), 
         class = gsub("D4", "", class))

melt_classes_mf38c <- melt_classes_mf38c %>%
  filter(D4 == "D4") %>%
  dplyr::select(cell, treatment, class, id, molpct_class_d4 = molpct_class) %>%
  inner_join(melt_classes_mf38c, .,by = join_by(id, cell, treatment, class)) %>%
  filter(D4 != "D4") %>%
  mutate(d4_pct = molpct_class_d4 / (molpct_class_d4 + molpct_class) * 100)


melt_classes_mf38c%>%
  filter(class %in% c("PE", "PEO-", "LPE", "LPEO-"),
         treatment == "ethD4")%>%
  ggplot(aes (x = cell, y = d4_pct, fill = class))+
  geom_bar(stat = "summary", fun = "geom_mean")+
  stat_summary(fun.data = geom_mean_sdl, geom= "errorbar", width = .4, linewidth = 0.2)+
  geom_point(size = 0.2, show.legend = F)+
  scale_y_continuous(expand = expansion( mult = c(0,0.025)))+
  scale_fill_manual(values = col_class)+
  facet_wrap(~class, nrow = 1)+
  labs(y = "2H4-labelled fraction (%)", x = NULL)+
  ggh4x::force_panelsizes(rows = unit(3, "cm"), cols =  unit(2, "cm")) 

ggsave(file.path(data_path, "Figures/pdf/Fig5k_length_db.pdf"), height = 2.5, width = 5)


```


# --- Supplementary figures ---

Fig S1 - WB, HeLa

## Fig S2 - HeLa, database filter

```{r}

filter_plot %>%
  ggplot(aes(x = lipid, y = db, fill = step))+
  geom_tile(linewidth = 0.5, color = "black")+
  scale_fill_manual(values = c('Remove by database' = "black", 
                               Included = "#003F7D", 
                               'Remove by blank' = "grey"))+
  scale_y_continuous(breaks = seq(0,100,1))+
  facet_wrap(~row, nrow = 2, scales = "free_x")+
  theme(strip.text = element_blank())+
  labs(y = "Total number of hydrocarbon chain bond", x = NULL , fill = NULL)+
  ggh4x::force_panelsizes(rows = unit(4, "cm"),
                          cols = unit(18, "cm"))


ggsave(file.path(data_path, "/Figures/pdf/Sup_FigS2_LipidDetectionAndFiltering.pdf"), height = 12, width = 25, units = "cm")

  
```

